\documentclass[sigconf,nonacm]{acmart}

%% Enable subfigures
\usepackage{subfigure}
%% Enable numbers in scientific format.
\usepackage{siunitx}
%% Enable enumerate start from.
\usepackage{enumitem}

%% Enable theorems
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

%% Enable algorithms
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\let\ReturnInline\Return
\renewcommand{\Return}{\State\ReturnInline}
\algrenewcommand\algorithmicrequire{$\rhd$}
\algrenewcommand\algorithmicensure{$\square$}

%% Fonts used in the template cannot be substituted; margin 
%% adjustments are not allowed.
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%% Title of the proceedings is different from ``Proceedings of ...''?
% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
% \acmPrice{15.00}
% \acmISBN{978-1-4503-XXXX-X/18/06}

%% Submission ID.
% \acmSubmissionID{123-A56-BU3}

%% Use the "author year" style of citations and references?
% \citestyle{acmauthoryear}

%% Message
\newcommand{\kk}[1]{{{\color{red} #1}}}
\newcommand{\ds}[1]{{{\color{blue} #1}}}
\newcommand{\su}[1]{{{\color{green} #1}}}

%% Ignore block
\newcommand{\ignore}[1]{}

%% Macros
\newcommand{\Lou}{\textit{Louvain}}
\newcommand{\LPA}{\textit{LPA}}
\newcommand{\Hyb}{\textit{Hybrid Louvain-LPA}}
\newcommand{\Sta}{\textit{Static}}
\newcommand{\Nai}{P-ND}
\newcommand{\DelOrg}{\textit{$\Delta$-screening}}
\newcommand{\Del}{P-DDS}
\newcommand{\Fro}{P-DF}
\newcommand{\StaLou}{\textit{Static Louvain}}
\newcommand{\NaiLou}{$\text{P-ND}_\text{L}$}
\newcommand{\DelLou}{$\text{P-DDS}_\text{L}$}
\newcommand{\FroLou}{$\text{P-DF}_\text{L}$}
\newcommand{\StaLPA}{\textit{Static LPA}}
\newcommand{\NaiLPA}{$\text{P-ND}_\text{LPA}$}
\newcommand{\DelLPA}{$\text{P-DDS}_\text{LPA}$}
\newcommand{\FroLPA}{$\text{P-DF}_\text{LPA}$}
\newcommand{\FroHyb}{$\text{P-DF}_\text{H}$}




\begin{document}

%% Full title of the paper.
\title[DF-Louvain: Fast Incrementally Expanding Approach for Community Detection on Dynamic Graphs]{DF-Louvain: Fast Incrementally Expanding Approach for Community Detection on Dynamic Graphs}

%% Short title to be used in page headers (optional).
% \title[short title]{full title}
% \subtitle{Something other than the title}

%% Authors and their affiliations.
\author{Subhajit Sahu}
\email{subhajit.sahu@research.iiit.ac.in}
\affiliation{%
  \institution{IIIT Hyderabad}
  \streetaddress{Professor CR Rao Rd, Gachibowli}
  \city{Hyderabad}
  \state{Telangana}
  \country{India}
  \postcode{500032}
}

%% Concise author list in page headers.
%\renewcommand{\shortauthors}{Sahu, Kothapalli, and Banerjee, et al.}

%% Show page numbers.
\settopmatter{printfolios=true}

%% Short summary of the work to be presented in the article.
\begin{abstract}
Community detection is the problem of recognizing natural divisions in networks. A relevant challenge in this problem is to find communities on rapidly evolving graphs. This calls for designing dynamic algorithms that reuse earlier output, only process vertices likely to change their community, extract parallelism by processing a batch of updates, and have low overhead.

This paper addresses the design of a high-speed community detection algorithm in the batch dynamic setting. First, our Dynamic Frontier (DF) approach is discussed. Given a batch update of edge deletion and insertions, this approach incrementally identifies an approximate set of affected vertices in the graph is with minimal overhead. We apply this approach to both Louvain, a high quality, and Label Propagation Algorithm (LPA), a high speed static community detection algorithm. Our approach achieves a mean speedup of $7.3\times$ and $6.7\times$ when applied to Louvain and LPA respectively, compared to our parallel and optimized implementation of $\Delta$-screening, a recently proposed state-of-the-art approach. Finally, we show how to combine Louvain and LPA with the Dynamic Frontier approach to arrive at a novel hybrid algorithm. This algorithm produces high-quality communities while providing a speedup of $2.0\times$ on top of DF Louvain.
\end{abstract}

%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10003809.10010170</concept_id>
<concept_desc>Theory of computation~Parallel algorithms</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003809.10003635</concept_id>
<concept_desc>Theory of computation~Graph algorithms analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

% \ccsdesc[500]{Theory of computation~Parallel algorithms}
% \ccsdesc[500]{Theory of computation~Graph algorithms analysis}

%% Pick words that accurately describe the work being presented.
\keywords{Community detection, Parallel Louvain algorithm}

% \received{20 February 2007}
% \received[revised]{12 March 2009}
% \received[accepted]{5 June 2009}



%% Process the author and title information.
\maketitle

\section{Introduction}
\label{sec:introduction}
\input{01-introduction}

\section{Related work}
\label{sec:related}
\input{02-related-work}

\section{Preliminaries}
\label{sec:preliminaries}
\input{03-preliminaries}

\section{Approach}
\label{sec:approach}
\input{04-approach}

\section{Evaluation}
\label{sec:evaluation}
\input{05-evaluation}

\section{Conclusion}
\label{sec:conclusion}
\input{06-conclusion}

%% The acknowledgments section.
\begin{acks}
I would like to thank Prof. Kishore Kothapalli,  Prof. Dip Sankar Banerjee, Balavarun Pedapudi, and Souvik Karfa for their support.
\end{acks}

%% Bibliography style to be used, and the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

\end{document}
\endinput
%% End of file.




%% NOTES:
%% - Parallelization seems to be not efficient for small batch updates.
%% - Discuss about conflicting updates
%% - 


%% TODO:
%% - Scale up the size of the graphs
%% - Move experiments to a better server
%% - Include a weak- and strong- scalabiilty plot: run the expt from 2 to 128 threads
%% - overall space planning
%% - add a few lines on novelty of the paper.
%% - table comparison of related work
%% - Include a section on preliminaries that talks about the various algorithmic ideas (Louvain, Label Propagation)

%% Workplan:
%% - KK -- Read Introduction, Related Work,
%% - Dip Sankar -- Approach -- summarize the main algorithmic ideas,
%% - Subhajit -- Results -- Plots, scalability, Dataset, experiments, implementation details,
